%\documentclass[10pt,preprint]{aastex}  % for e-submission to ApJ
%\documentclass[12pt,preprint2]{aastex}  % for e-submission to ApJ - two column
\documentclass[onecolumn]{emulateapj}  % this makes everything look like ApJ

\usepackage{graphicx, natbib, color}

%%%% PUT NEW COMMANDS AND DEFINITIONS HERE %%%%
\input{local-commands}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Paper Outline  (last updated June 15, 2009)
%
% 1.  Introduction [GB,MN]
% 2.  Overview [GB]
% - including both equations and a few paragraph overview of how each method works
% - this might have section 2.1, 2.2, etc. corresponding to very brief overviews of section 3.1, 3.2, etc.
% 3.  Numerical methods
%  3.1 AMR [GB]
%  3.2. Hydro: ZEUS [GB]
%  3.3. Hydro: PPM [GB,AK]
%  3.4. Gravity [GB]
%  3.5. Particle Integration [GB]
%  3.6. Chemistry [BWO]
%  3.7. Star formation & Feedback [BWO]
%  3.8. Timestep constraints [BWO,GB]
% 4. Software Implementation
% - maybe this could be split up by into sections 4.1, 4.2, 4.3, etc. using the same scheme as section 3 [same responsibilities as 3]
% 5. Parallelism 
% 5.1. Parallel strategy [GB, RH]
% 5.2. Parallel performance [JB, GB, RH]
% 6. Code Tests [All]
% 7. Summary and Conclusions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{aa}  % correct formatting for ApJ style files

\begin{document}

\title{Enzo: An Adaptive Mesh Refinement Code for Astrophysics}
\author{Greg Bryan\altaffilmark{1}, Michael L. Norman\altaffilmark{2}, 
Brian W. O'Shea\altaffilmark{3}, Robert Harkness\altaffilmark{4}, Tom Abel\altaffilmark{5},
David C. Collins,\altaffilmark{2} 
Alexei Kritsuk\altaffilmark{2} and James Bordner\altaffilmark{2} }

\altaffiltext{1}{ Department of Astronomy, Columbia University,
New York, NY, 10027, U.S.A.  Email: gbryan@astro.columbia.edu }
\altaffiltext{2}{Center for Astrophysics and Space Sciences,
University of California at San Diego.
Email: mnorman,dcollins,akritsuk,jbordner@cosmos.ucsd.edu}
\altaffiltext{3}{Theoretical Astrophysics (T-6), Los Alamos National
Laboratories.  Email:  bwoshea@lanl.gov}
\altaffiltext{4}{San Diego Supercomputing Center, University of California
at San Diego.  Email: harkness@sdsc.edu}
\altaffiltext{5}{Kavli Institute for Particle Astrophysics and Cosmology,
 Stanford University.  Email:  tabel@stanford.edu}

\begin{abstract}
We describe the \enzo\ code, which uses a structured adaptive mesh refinement technique to provide high spatial and temporal resolution for modeling astrophysical flows.  The code includes support for the following physics: multi-dimensional hydrodynamics, self-gravity, N-body dynamics, primordial gas chemistry, optically-thin radiative cooling, and cosmological expansion.  Two Eulerian hydrodynamic solvers are provided.  This paper describes the method as well as the code's performance on a wide range of test problems.
\end{abstract}

\keywords{methods: numerical --- hydrodynamics}

\maketitle

%% Section: introduction
\input{introduction}

%% Section:  Overview
\input{overview}

%% Section: numerical methods
\section{Numerical Methods}
\label{sec.methods}

We now describe each of the numerical methods in detail.  The structure of this section mirrors the previous one, in which each method was briefly described; here we present the details.

\input{numerical-amr}
\input{numerical-ppm}
\input{numerical-zeus}
\input{numerical-gravity}
\input{numerical-particle}
\input{numerical-chemistry}
\input{numerical-starformation}
\input{numerical-timestep}
%\input{numerical-units}

%% Section: software implementation
% \input{software}

%% Section: code tests
%%  this should probably be done with one file per test
\input{code-tests}

%% Section: parallelism and performance
\input{parallel}

%% Section: discussion
\input{discussion}

%% Section: summary
\input{summary}

\appendix

%% Appendix: interpolation
%\input{app-interpolation}

%% Appendix:  Data Structures
\input{app-datastruct}

%% Appendix:  I/O
\input{app-IO}

%% Appendix:  Zeus hydro
%\input{app-zeushydro}

%% Appendix:  PPM hydro
%\input{app-ppmhydro}

%% Appendix:  gravity solver
%\input{app-gravity}

%% Appendix:  
\input{app-chemistry}

\bibliographystyle{apj}
\bibliography{apj-jour,ms}  % looks in ms.bib for bibliography info

\end{document}  


